From Reversible Programs to \\ Univalent Universes and Back

### 2021-07-02 Pi+/paper/finite.tex
Move the definition of Sn

A group presentation, as defined in~\label{def:presentation}, requires the relation to be on the set of words $A + A$,
where the right copy corresponds to the set of formal inverses of the generators. We will resolve this mismatch later
in~\cref{prop:lift-codiag}.

\todo{check: However, in our case, the constructor $\cancel$ specifies that the inverse of each element is again the
  same element.}

@@

By strong normalisation, we get a unique choice function $\normf : {\List[\Fin[n]]} \to {\List[\Fin[n]]}$ that produces
a normal form for terms of $\List[\Fin[n]]$. Equivalence classes on $\List[\Fin[n]]$ are defined to be those terms that
have the same normal form. We state a few important properties that $\normf$ satisfies.

\begin{proposition}
  \leavevmode
  \begin{enumerate}
    \item For all $l : \List[\Fin[n]]$, we have that $l \cox* \normf(l)$.
    \item $\normf$ is idempotent, that is, $\normf \comp \normf \htpy \normf$.
    \item $\normf$ splits into a section-retraction pair, that is, we have
          ${\List[\Fin[n]]} \xrightarrow{s} \Sn[n] \xrightarrow{r} {\List[\Fin[n]]}$ such that $s \comp r \htpy \normf$
          and $r \comp s \htpy \idfunc_{\Sn[n]}$.
    \item \(\im{\quotinc} \eqv \Sn \eqv \im{\normf} \).
  \end{enumerate}
\end{proposition}

\todo{We set out to solve the word problem for $\Sn$ \ldots}
\todo{Revise this section by introducing $\Sn$ here.}

Note that this relation $\cox*$ is not $\hProp$-valued, that is, $w \cox* v$ is not a proposition because reductions are
not unique. Because of this, the quotient $\Sn$ is not effective, that is, $\quotrel : w \cox* v \to q(w) \id q(v)$ is
not an equivalence. We could truncate the relation to change this, but then the codomain of $\normf$ also needs to be
truncated. Instead, using the $\normf$ function, we could define a new relation
$(w \approx v) \defeq (\normf(w) \id \normf(v))$ which is $\hProp$-valued, and quotient $\List[\Fin[n]]$ by $\approx$.
This is not necessary because our results hold for $\cox*$ using the properties of the rewriting system we proved, and
we do not rely on the effectivity of the quotient.

Finally, we prove the group structure of $\Sn$, and show that it is indeed a group presentation, given by a quotient of
the free group $F(\Fin[n])$.

\begin{proposition}
  There is a group structure on $\Sn$, where the identity element is $\nil$, multiplication is given by list append, and
  inverse is given by list reversal.
\end{proposition}

\begin{proposition}~\label{prop:lift-codiag}
  $\Sn$ is equivalent to the generated group given by the normal closure of $\cox*$ extended to
  $\List(\Fin[n] + \Fin[n])$ along the codiagonal map $\nabla_{A} : A + A \to A$.
\end{proposition}

>>>>

@@

By strong normalisation, we get a unique choice function $\normf : {\List[\Fin[n]]} \to {\List[\Fin[n]]}$ that produces
a normal form for terms of $\List[\Fin[n]]$. We state and prove a couple of important properties enjoyed by $\normf$.

\begin{proposition}
  \leavevmode
  \begin{enumerate}
    \item For all $l : \List[\Fin[n]]$, we have that $l \cox* \normf(l)$.
    \item $\normf$ is idempotent, that is, $\normf \comp \normf \htpy \normf$.
  \end{enumerate}
\end{proposition}

Note that this relation $\cox*$ is not $\hProp$-valued, that is, $w \cox* v$ is not a proposition because reductions are
not unique. Because of this, the quotient $\Sn$ is not effective, that is, $\quotrel : w \cox* v \to q(w) \id q(v)$ is
not an equivalence. We could truncate the relation to change this, but then the codomain of $\normf$ also needs to be
truncated.

We could also define equivalence classes on $\List[\Fin[n]]$ to be those terms that have the same normal form. Using the
$\normf$ function, we could define a new relation $(w \approx v) \defeq (\normf(w) \id \normf(v))$ which is
$\hProp$-valued, and quotient $\List[\Fin[n]]$ by $\approx$. Using the properties of the rewriting system we proved, we
make a few observations.

\begin{proposition}
  \leavevmode
  \begin{enumerate}
    \item $\normf$ splits into a section-retraction pair, that is, we have ${\List[\Fin[n]]} \xrightarrow{s} \Sn[n]
            \xrightarrow{r} {\List[\Fin[n]]}$ such that $s \comp r \htpy \normf$ and $r \comp s \htpy \idfunc_{\Sn[n]}$.
    \item \(\im{\quotinc} \eqv \Sn \eqv \im{\normf} \).
  \end{enumerate}
\end{proposition}

Notice however, that a group presentation, as defined in~\cref{def:presentation}, requires the relation to be on the set
of words $A + A$, where the right copy corresponds to the set of formal inverses of the generators. The constructor
$\cancel$ specifies that the inverse of each element is again the same element, using which we can show that our
definition of $\Sn$ is equivalent to the definition of a presented group, by lifting the $\cox*$ relation.

\begin{proposition}
  \leavevmode
  \begin{enumerate}
    \item There is a group structure on $\Sn$, where the identity element is $\nil$, multiplication is given by list
          append, and inverse is given by list reversal.
    \item $\Sn$ is equivalent to the group presented by generators $\Fin[n] + \Fin[n]$ with the relations given by the
          normal closure of $\cox*$ extended to $\List(\Fin[n] + \Fin[n])$ along the codiagonal map $\nabla_{A} : A + A
          \to A$.
  \end{enumerate}
\end{proposition}

We set out to solve the word problem for $\Sn$. To decide if two words in $\List[\Fin[n]]$ are equal, we simply have to
compute their normal forms using $\normf$ and check if they're equal on the nose. If they do, they correspond to the
same permutation.
